---
title: "Notation for FDA"
output:
  html_document: 
    code_folding: hide
    toc: true
    toc_float: true
---

```{r setup, echo = TRUE, show = "hide", message=FALSE, warning=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "##",
  fig.width = 8, 
  fig.height = 5.5,
  out.width = "90%"
)

library(tidyverse)
library(viridisLite)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

This page contains a combination of traditional lecture materials (slides) and code demonstrating the relevant methods. The short course will proceed by working through both. We will use several recent packages in our examples; see the [About](./About.html) page for information about the package versions.

```{r}
library(tidyverse)
library(refund.shiny)
library(tidyfun)
```

## Notation for functional data

<script async class="speakerdeck-embed" data-id="22db63f1cf324ec6b998c631191af6ed" data-ratio="1.77777777777778" src="//speakerdeck.com/assets/embed.js"></script>
<div style="margin-bottom:5px"> <strong> <a href="https://speakerdeck.com/jeffgoldsmith/short-course-notation" title="Wearables Short Course: Notation for FDA" target="_blank">Wearables Short Course: Notation for FDA</a> </strong> from <strong><a href="https://speakerdeck.com/jeffgoldsmith" target="_blank">Jeff Goldsmith</a></strong>. </div>

Slides can be downloaded [here](./SlidesPPT/ShortCourse_Notation.pdf).

## Practical example

In this section we will use the [HeadStart data](./shortcourse_data.html) as an example to review notation and demonstrate useful plots of functional data. The code below loads this dataset.

```{r headstart_load, eval = TRUE}

load("./DataCode/HeadStart.RDA")

```

We're also going to use the [`tidyfun`](https://tidyfun.github.io/tidyfun/index.html) R package for some plotting and exploratory analysis. 

First, we plot a single subject to illustrate the "functional" nature of these data. 

```{r headstart_onesubj, eval = TRUE}
covariate_data = 
  covariate_data %>% 
  mutate(accel = tfd(accel))

covariate_data %>% 
	filter(id == 1) %>%
	ggplot(aes(y = accel)) + 
  geom_spaghetti()
```

Next, we plot all subjects. The previous plot is a single noodle in this plot of spaghetti.

```{r headstart_spaghetti, eval = TRUE}
covariate_data %>% 
	ggplot(aes(y = accel)) + 
  geom_spaghetti(alpha = .1)
```

Next we'll look at some raw and smoothed mean functions. 

```{r headstart_mean, eval = TRUE}
covariate_data %>% 
  group_by(season, sex) %>% 
  summarize(mean_act = mean(accel)) %>%
  mutate(smooth_mean = tf_smooth(mean_act)) %>%
  ggplot(aes(y = mean_act, color = sex)) +
  geom_spaghetti(alpha = .5) +
  geom_spaghetti(aes(y = smooth_mean), size = 2) + 
  facet_grid(~season)
```

Including observed data is also possible, and helps to provide context for the mean values.

```{r headstart_smooth_mean, eval = TRUE}
mean_df = 
  covariate_data %>% 
  group_by(season, sex) %>% 
  summarize(mean_act = mean(accel)) %>%
  mutate(smooth_mean = tf_smooth(mean_act))

covariate_data %>% 
  ggplot(aes(y = accel, color = sex)) +
  geom_spaghetti(alpha = .05) +
  geom_spaghetti(data = mean_df, aes(y = smooth_mean), size = 2) + 
  facet_grid(~season)
```

Rainbow plots color each noodle according to some covariate value. We illustrate this using BMI Z-score. 

```{r headstart_rainbow, eval = TRUE}
covariate_data %>% 
  ggplot(aes(y = accel, color = BMIZ)) +
  geom_spaghetti(alpha = .05)
```

Lastly, we show the covariance surface to get an idea of the overall variance and the correlation across times. 

```{r headstart_covariance, eval = TRUE, fig.height = 5, fig.width = 7}

library(plotly)

covariance = cov(accel)
plot_ly(z = ~covariance) %>% add_surface()

```
